Msprime 1.0: an efficient and extensible coalescent simulation framework

### 2021-07-22 paper.tex
Finalise msprime benchmark.

Even when we consider a single tree, \msprime\ is far more
efficient than methods such as \SeqGen\ and \Pyvolve,
and has similar performance to \phastsim.

Comparing the performance of mutation generation methods is problematic.
Because there is no standard format for representing trees and
the accompanying mutations, all existing methods are focused on
writing out the resulting sequences. Methods are also not modular,
and so it is very difficult to disentangle the time required to
read in the tree file and output the sequences from the actual
time required to perform the simulation.

\jkcomment{We probably want to keep a version of this narrative too.}
For example, we simulated mutations under the JC69
model for a tree with $10^6$ leaves over $10^4$ sites (resulting
in $\sim 280000$ mutations) in less that 1.5 seconds, including
the time required for file input and output. While a systematic
comparison of run-times is not feasible here, this is orders
of magnitude faster than methods such as
Seq-Gen~\citep{rambaut1997seq} and
Pyvolve~\citep{spielman2015pyvolve}
and at least competitive with the recent method from
\cite{demaio2021phastsim}.

\textbf{TODO: would love to do a like-for-like comparison, but quite hard}

>>>>

% python3 evaluation/generate_mutations_perf_data.py benchmark-single-tree
Even if we consider a single tree, \msprime\ generates mutations
extremely efficiently.
For example, we simulated mutations under the BLOSUM62 amino
acid model for a tree with $10^6$ leaves over $10^4$ sites (resulting
in $\sim 260,000$ mutations) in about $0.8$ seconds, including
the time required for file input and output.
We do not attempt a systematic benchmarking of \msprime's
mutation generation code against other methods because it is
difficult to disentangle the effects of inefficient input and
output formats from the mutation generation algorithms
at this scale. It seems quite unlikely, however,
that generating mutations with \msprime\ would be a bottleneck in any
realistic analysis.
